### 2013-06-03 papers/bprheo_r/bprheo.tex
Updated all pictures for bprheo paper
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@1994 40876441-3496-0410-a613-a2fca35b2ae2

@@

If the flow rate exceeds a certain critical value BPII does not undergo the 
homogeneous deformation described above and adopts another mode of flow. 
Fig. \ref{bp2-high} shows a sequence of cuts through the director 
field at high shear rate. The entire blue phase network has broken up 
into a simple cholesteric helix and flows as a travelling 
helical wave with the helical axis oriented along the vorticity direction.
The state is translationally invariant in the flow and gradient directions.
It is the preferred mode of flow of a standard cholesteric phase
in vorticity mode at low Ericksen numbers. This is followed by
the uncoiling transition to a nematic, flow-aligned state at even 
higher shear rates ~\cite{Rey:1996a, Rey:1996b}.

@@

\begin{figure}[htpb]
\includegraphics[width=0.495\textwidth]{dir+y-250k_run949.png}
\includegraphics[width=0.495\textwidth]{dir+y-253k_run949.png}
\includegraphics[width=0.495\textwidth]{dir+y-255k_run949.png}
\includegraphics[width=0.495\textwidth]{dir+y-257k_run949.png}
\caption{Director field of BPII at high shear rate: The images show snapshots at time steps 
$t=2.50, 2.53,2.55, 2.57\e{5}$ (from top to bottom) with flow into/out of the page. The colour code gives the alignment of the director field in the flow direction (x) with red fully aligned and blue out of plane. 
The configuration is a helical wave that travels along the direction of vorticity.
It is translationally invariant along the gradient and flow direction.}
\label{bp2-high}
\end{figure}

@@

\begin{figure*}[htpb]
\includegraphics[width=0.245\textwidth]{disc-xy-10k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-180k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-200k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-210k_run1115.png}\\
\includegraphics[width=0.245\textwidth]{disc-xy-400k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-700k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-750k_run1115.png}
\includegraphics[width=0.245\textwidth]{disc-xy-2200k_run1115.png}
\caption{Snapshots of BPI disclination network at $\dot{\gamma}=4.88\e{-6}$: 
Depicted is the transition from the quiescent state to flow-induced, intertwined 
helices that undergo a recurring structural transformation. The top row
shows a section of one unit cell for early times 
$t=1\e{4}, 1.8\e{5}, 2.0\e{5}$ and $2.1\e{5}$. The bottom 
pictures on the far left, centre left and centre right 
show the situation at later time steps $t=4.0, 7.0$ and $7.5\e{5}$,
which the system passes for several cycles before it ends up in
an amorphous state (bottom row far right, at time step $t=2.2\e{6}$).}
\label{bp1-low}
\end{figure*}

\if{The negative branch in the shear stress is connected
to a local maximum followed by a local minimum in the 
average free energy. Hence, the region of negative stress can be 
regarded as a phase during which elastic forces exert a 
'pull' on the network while it tries to reach a more 
favourable configuration. However, the low flow rate
causes this unstable intermediate state to be exposed to these
elastic forces for a relatively long time, long enough to 
create small distortions which finally break the symmetry of the
flow-induced configuration.}\fi

@@

\if{In order to gain a better understanding of the complex oscillatory behaviour in regime BPI-2 
we performed a Fourier analysis of the shear stress $\Pi_{xy}$ according to Eq. \ref{spectrum}. 
The spectrum is shown in Fig. \ref{bp1-spectrum}.

@@

The intricacy and periodic recurrence of the flow-induced conformations 
in BPI raises the question: how robust are these results to variations in 
system size? Although a pitch length of 64 LBU provides enough resolution 
for tracking the network in flow it is not clear whether $4^3=64$ unit cells 
in total are actually sufficient to avoid finite size effects.
Unfortunately the computational costs of larger simulations are still
quite prohibitive (in the region of ${\it O}(10^5)$ core hours per run at
doubled system size), given that large strains have to
be simulated in order to reach a steady state. Therefore we opted for a 
comparison with smaller runs.

\begin{figure}[htpb]
\includegraphics[width=0.43\textwidth]{disc+y-600k-run911_run1163.png}\\
\includegraphics[width=0.495\textwidth]{stress_bp1_2uc_4uc.pdf}
\caption{Disclination network and shear stress $\Pi_{xy}$ in BPI: 
The dark (red) network in top picture shows a snapshot of run that consists 
of $2^3=8$ unit cells, whereas the light (cyan) network displays the 
central section of 8 unit cells in a bigger run with $4^3=64$ unit cells altogether. 
The bottom picture shows the shear stress over time. (Interestingly, 
an intermediate sized run with $3^3=27$ unit cells led to results that 
were virtually identical to the smaller run with 8 unit cells.) 
The shear rate was $\gd=7.81\e{-5}$.
}
\label{bp1-2uc4uc}
\end{figure}

Fig. \ref{bp1-2uc4uc} shows a comparison of the disclination network and 
shear stress $\Pi_{xy}$ of two different runs, one consisting of 8 
unit cells and the other one with 64 unit cells.
We observe that there are deviations in the visual appearance of the 
network as well as in the time-dependence of the shear stress.
Perhaps surprisingly, the smaller run with 8 unit cells takes longer 
to reach the steady state with its flow-induced recurring disclination network. 
These results indicate that we cannot know with certainty whether the individual 
flow-induced conformations that we find for 64 unit cells are affected 
by the imposed periodicity -- quite possibly the details of the oscillations
would differ in an even larger simulation. 
Nevertheless, the comparison suggests that 
the recurrence period and the complexity of the network in flow are  
generic features of BPI in this regime of flow.

@@

At high flow rates, $6.25\e{-4}\lesssim\gd\lesssim7.81\e{-4}$ ($10.67\lesssim {\it Er}\lesssim 13.33$), 
we observe a break-up of the flow-induced 
disclination network of regime BPI-2. However, 
in contrast to BPII, where the configuration is a travelling helical 
wave (cf. Fig. \ref{bp2-high}), BPI breaks down into a quasi 
two-dimensional, translationally invariant state, which is shown in 
Fig. \ref{bp1-high}. 
There are distinctive regions where the director field is 
predominantly oriented along the vorticity direction and 
other regions where it is oriented rather in flow-gradient plane. 
Occasionally, while moving past each other, these rolls resemble 
regions of local double twist.
Interestingly the critical Ericksen number for the break-up lies in 
about the same window as that of transition between BPII-1 and BPII-2.

\begin{figure}[htpb]
\includegraphics[width=0.45\textwidth]{dir3d-z-302k_run916.png}
\caption{Director field of BPI at high shear rate: The pictures show a region 
of $L_x\times L_y \times L_z= 32^3$ lattice sites with the flow/gradient 
direction oriented along the horizontal/vertical axis. 
The colour code gives the magnitude of the z-component of the director field. 
The data show that the structure is quasi-2D, with rolling double twist
patterns in the $xy$ plane, and an approximate translational symmetry 
along the vorticity direction.}
\label{bp1-high}
\end{figure}

@@

\begin{figure}[htpb]
\includegraphics[width=0.45\textwidth]{dir3d+y-200k_run1179.png}
\caption{Flow induced cholesteric state. The orientation of the helix is 
along the gradient direction. The director field is quasi-static during 
the flow, which occurs in ``nematic layers''. Colour coding indicates the 
$x$ component of the director field (see colorbar on the right).
For clarity only selected sites are shown along the $y$-direction (along which
the structure is translationally invariant).}
\label{fig:cholflow}
\end{figure}

@@

% If you have acknowledgments, this puts in the proper section head.
\section*{Acknowledgments}
We acknowledge support by EPSRC grant nos. EP/E045316 and EP/E030173, 
the MAPPER EU-FP7 project (grant no. RI-261507) and computing time on HECToR.
This work was granted access to the HPC resource of CSC, Finland, made available 
within DECI by the PRACE-2IP, funded by the European Community's 7th Framework 
Program under grand agreement No. RI-283493. We thank Peter J. Collings 
for very stimulating discussions. 
MEC holds a Royal Society Research Professorship.

>>>>

\newif\ifold
\oldtrue

@@

% New conformation is presumably metastable %
\ifold{If the flow rate exceeds a certain critical value BPII does not undergo the 
homogeneous deformation described above and adopts another mode of flow. 
Fig. \ref{bp2-high} shows a sequence of cuts through the director 
field at high shear rate. The entire blue phase network has broken up 
into a simple cholesteric helix and flows as a travelling 
helical wave with the helical axis oriented along the vorticity direction.
The state is translationally invariant in the flow and gradient directions.
It is the preferred mode of flow of a standard cholesteric phase
in vorticity mode at low Ericksen numbers. This is followed by
the uncoiling transition to a nematic, flow-aligned state at even 
higher shear rates ~\cite{Rey:1996a, Rey:1996b}.

@@

\begin{figure}[htpb]
%\includegraphics[width=0.495\textwidth]{dir+y-250k_run949.png}
%\includegraphics[width=0.495\textwidth]{dir+y-253k_run949.png}
%\includegraphics[width=0.495\textwidth]{dir+y-255k_run949.png}
%\includegraphics[width=0.495\textwidth]{dir+y-257k_run949.png}
\caption{Director field of BPII at high shear rate: The images show snapshots at time steps 
$t=2.50, 2.53,2.55, 2.57\e{5}$ (from top to bottom) with flow into/out of the page. The colour code gives the alignment of the director field in the flow direction (x) with red fully aligned and blue out of plane. 
The configuration is a helical wave that travels along the direction of vorticity.
It is translationally invariant along the gradient and flow direction.}
\label{bp2-high}
\end{figure}
\fi

@@

\begin{figure*}[htpb]
\includegraphics[width=0.245\textwidth]{disc-xy-10k_run1115r.png}
\includegraphics[width=0.245\textwidth]{disc-xy-180k_run1115r.png}
\includegraphics[width=0.245\textwidth]{disc-xy-200k_run1115r.png}
\includegraphics[width=0.245\textwidth]{disc-xy-210k_run1115r.png}\\
\includegraphics[width=0.245\textwidth]{disc-xy-1650k_run1115r.png}
\includegraphics[width=0.245\textwidth]{disc-xy-1800k_run1115r.png}
\includegraphics[width=0.245\textwidth]{disc-xy-2000k_run1115r.png}
\includegraphics[width=0.245\textwidth]{disc-xy-2400k_run1115r.png}
\caption{Snapshots of BPI disclination network at $\dot{\gamma}=4.88\e{-6}$: 
Depicted is the transition from the quiescent state to flow-induced, intertwined 
helices that undergo a recurring structural transformation. The top row
shows a section of one unit cell for early times 
$t=1\e{4}, 1.8\e{5}, 2.0\e{5}$ and $2.1\e{5}$. The bottom 
pictures on the far left, centre left and centre right 
show the situation at later time steps $t=1.65, 1.8$ and $2.0\e{6}$,
which the system passes for several cycles before it ends up in
an amorphous state (bottom row far right, at time step $t=2.4\e{6}$).}
\label{bp1-low}
\end{figure*}

\ifold{The negative branch in the shear stress is connected
to a local maximum followed by a local minimum in the 
average free energy. Hence, the region of negative stress can be 
regarded as a phase during which elastic forces exert a 
'pull' on the network while it tries to reach a more 
favourable configuration. However, the low flow rate
causes this unstable intermediate state to be exposed to these
elastic forces for a relatively long time, long enough to 
create small distortions which finally break the symmetry of the
flow-induced configuration.}\fi

@@

\ifold{In order to gain a better understanding of the complex oscillatory behaviour in regime BPI-2 
we performed a Fourier analysis of the shear stress $\Pi_{xy}$ according to Eq. \ref{spectrum}. 
The spectrum is shown in Fig. \ref{bp1-spectrum}.

@@

% No differences in retake %
\ifold{The intricacy and periodic recurrence of the flow-induced conformations 
in BPI raises the question: how robust are these results to variations in 
system size? Although a pitch length of 64 LBU provides enough resolution 
for tracking the network in flow it is not clear whether $4^3=64$ unit cells 
in total are actually sufficient to avoid finite size effects.
Unfortunately the computational costs of larger simulations are still
quite prohibitive (in the region of ${\it O}(10^5)$ core hours per run at
doubled system size), given that large strains have to
be simulated in order to reach a steady state. Therefore we opted for a 
comparison with smaller runs.

\begin{figure}[htpb]
%\includegraphics[width=0.43\textwidth]{disc+y-600k-run911_run1163.png}\\
%\includegraphics[width=0.495\textwidth]{stress_bp1_2uc_4uc.pdf}
\caption{Disclination network and shear stress $\Pi_{xy}$ in BPI: 
The dark (red) network in top picture shows a snapshot of run that consists 
of $2^3=8$ unit cells, whereas the light (cyan) network displays the 
central section of 8 unit cells in a bigger run with $4^3=64$ unit cells altogether. 
The bottom picture shows the shear stress over time. (Interestingly, 
an intermediate sized run with $3^3=27$ unit cells led to results that 
were virtually identical to the smaller run with 8 unit cells.) 
The shear rate was $\gd=7.81\e{-5}$.
}
\label{bp1-2uc4uc}
\end{figure}

Fig. \ref{bp1-2uc4uc} shows a comparison of the disclination network and 
shear stress $\Pi_{xy}$ of two different runs, one consisting of 8 
unit cells and the other one with 64 unit cells.
We observe that there are deviations in the visual appearance of the 
network as well as in the time-dependence of the shear stress.
Perhaps surprisingly, the smaller run with 8 unit cells takes longer 
to reach the steady state with its flow-induced recurring disclination network. 
These results indicate that we cannot know with certainty whether the individual 
flow-induced conformations that we find for 64 unit cells are affected 
by the imposed periodicity -- quite possibly the details of the oscillations
would differ in an even larger simulation. 
Nevertheless, the comparison suggests that 
the recurrence period and the complexity of the network in flow are  
generic features of BPI in this regime of flow.
\fi

@@

\ifold At high flow rates, $6.25\e{-4}\lesssim\gd\lesssim7.81\e{-4}$ ($10.67\lesssim {\it Er}\lesssim 13.33$), 
we observe a break-up of the flow-induced 
disclination network of regime BPI-2. However, 
in contrast to BPII, where the configuration is a travelling helical 
wave (cf. Fig. \ref{bp2-high}), BPI breaks down into a quasi 
two-dimensional, translationally invariant state, which is shown in 
Fig. \ref{bp1-high}. 
There are distinctive regions where the director field is 
predominantly oriented along the vorticity direction and 
other regions where it is oriented rather in flow-gradient plane. 
Occasionally, while moving past each other, these rolls resemble 
regions of local double twist.
Interestingly the critical Ericksen number for the break-up lies in 
about the same window as that of transition between BPII-1 and BPII-2.

\begin{figure}[htpb]
%\includegraphics[width=0.45\textwidth]{dir3d-z-302k_run916.png}
\caption{Director field of BPI at high shear rate: The pictures show a region 
of $L_x\times L_y \times L_z= 32^3$ lattice sites with the flow/gradient 
direction oriented along the horizontal/vertical axis. 
The colour code gives the magnitude of the z-component of the director field. 
The data show that the structure is quasi-2D, with rolling double twist
patterns in the $xy$ plane, and an approximate translational symmetry 
along the vorticity direction.}
\label{bp1-high}
\end{figure}
\fi

@@

\begin{figure}[htpb]
\includegraphics[width=0.475\textwidth]{dir3d+y-200k_run1179r.png}
\caption{Flow induced cholesteric state. The orientation of the helix is 
along the gradient direction. The director field is quasi-static during 
the flow, which occurs in ``nematic layers''. Colour coding indicates the 
$x$ component of the director field (see colorbar on the right).
For clarity only selected sites are shown along the $y$-direction (along which
the structure is translationally invariant).}
\label{fig:cholflow}
\end{figure}

@@

% If you have acknowledgments, this puts in the proper section head.
\section*{Acknowledgments}

This work was granted access to the HPC resource of CSC, Finland, made available 
within DECI by the PRACE-2IP, funded by the European Community's 7th Framework 
Program under grand agreement No. RI-283493. We also acknowledge support by 
EPSRC grant nos. EP/E045316 and EP/E030173 and the MAPPER EU-FP7 project 
(grant no. RI-261507).We thank Peter J. Collings for stimulating discussions. 
MEC holds a Royal Society Research Professorship.
